An AI is Haunting Campus^[Karl Marx wrote about a \"spectre...haunting Europe--the spectre of communism\"; GenAI looks awfully like the spectre of unregulated surveillance capitalism.]: A Call for Human-Centered Higher Education

### 2024-10-28 paper/paper.qmd
Submission to Higher Education

>>>>

## Abstract

The use of Generative Artificial Intelligence (AI) in higher education has been a sensational topic since the launch of ChatGPT and similar tools. AI has been presented as an everything tool that can do a fabulous array of tasks, from creating content to grading to providing feedback to students. This presentation of AI is misleading, and there is a need to examine the claims about what the technology can do and what it actually does in the context of teaching and learning. In this paper I scan the academic and gray literature, along with media reports regarding AI, to show how discussions about AI hinge on making a large number of unverified claims (Gish Gallop) while presenting the technology as inevitable (a False Dilemma). However, once we examine what AI does, and particularly what it does for whom and to whom, a picture emerges of a technology that is intrinsically inimical to the principles that most institutions of higher education claim to hold dear, including the best interest of their students.
